|

### 2021-08-11 prova_knit_finale.Rmd
Final commit

@@

In the following sections, we will try to replicate some of the analyses which were previously conducted on synthetic data, generated by a basic stochastic volatility model; this time, the same model specification and filters will be used on real data. However, as we will see, real data require different strategies to evaluate the effectiveness of several Sequential Monte Carlo techniques. Intuitively, such differences are due to the fact that now the latent process driving the Hidden Markov Model is not observed, nor do we know exactly its law of motion.\
This part of the analysis is structured as follows: section \ref{Literary review} provides a review of recent research in the stochastic volatility literature, including extensions of the basic specification and the use of particle filters; section \ref{Data description} describes the data which we employ and the transformations that were applied to them. In section \ref{Comparing the filters} we briefly compare the series of filtered states that are obtained from different particle filters; moreover, several PFs are compared in terms of computational cost in approximating the exact filtering distribution implied by our SV model. Section \ref{Forecast comparison} features a forecast comparison between a SV specification and a model that assumes constant mean and variance for the financial returns, namely a forecasting exercise that would not be possible were it not for particle filtering techniques. In section \ref{Realised volatility} we conclude by employing a non-parametric measure of realised volatility as a proxy for the latent volatility process, allowing a RMSE and MAE comparison between filters which is similar to that of section \ref{Simulated SV data} for synthetic data.

@@

```{r cll_plot, echo=FALSE, fig.align='center', fig.cap="Relative log-likelihood of S&P500 daily log returns, CV and SV", fig.height=3, fig.pos='H', fig.width=10, results=F}
load("SV/cll_plot.Rda")
cll_plot+
  theme_bw()+
  scale_x_date(date_breaks = "3 months",date_minor_breaks="1 month",date_labels = "%m/%y")+
  labs(x="Time (month/year)",y="")
```
Overall, this analysis shows how a model with stochastic
volatility is better at describing and predicting the considered time-frame of S&P500 returns. As can be seen from the graph and perhaps unsurprisingly,
the density forecast given the stochastic volatility model represents
a particularly significant improvement in the period of extreme values of returns in March 2020.

>>>>

where by the inequality sign we mean that the matrix $Var\big[\psi^{A_t^{n}}\big]$ is positive definite.

@@

In the following sections, we will try to replicate some of the analyses which were previously conducted on synthetic data, generated by a basic stochastic volatility model; this time, the same model specification and filters will be used on real data. However, as we will see, real data require different strategies to evaluate the effectiveness of several Sequential Monte Carlo techniques. Intuitively, such differences are due to the fact that now the latent process driving the Hidden Markov Model is not observed, nor do we know exactly its law of motion.\
This part of the analysis is structured as follows: section \ref{Literary review} provides a review of recent research in the stochastic volatility literature, including extensions of the basic specification and the use of particle filters; section \ref{Data description} describes the data which we employ and the transformations that were applied to them. In section \ref{Comparing the filters} we briefly compare the series of filtered states that are obtained from different particle filters; moreover, several PFs are compared in terms of computational cost in approximating the exact filtering distribution implied by our SV model. Section \ref{Realised volatility} employs a non-parametric measure of realised volatility as a proxy for the latent volatility process, allowing a RMSE and MAE comparison between filters which is similar to that of section \ref{Simulated SV data} for synthetic data. In section \ref{Forecast comparison} we conclude by providing a forecast comparison between a SV specification and a model that assumes constant mean and variance for the financial returns, namely a forecasting exercise that would not be possible were it not for particle filtering techniques.

@@

```{r cll_plot, echo=FALSE, fig.align='center', fig.cap="Relative log-likelihood of SP500 daily log returns, CV and SV", fig.height=3, fig.pos='H', fig.width=10, results=F}
load("SV/cll_plot.Rda")
cll_plot+
  theme_bw()+
  scale_x_date(date_breaks = "3 months",date_minor_breaks="1 month",date_labels = "%m/%y")+
  labs(x="Time (month/year)",y="")
```
Overall, this analysis shows how a model with stochastic
volatility is better at describing and predicting the considered time-frame of S&P500 returns. As can be seen from the graph and perhaps unsurprisingly,
the density forecast given the stochastic volatility model represents
a particularly significant improvement in the period of extreme values of returns in March 2020.
